---
title: "Mercari Price Suggestion Challenge"
author: "Ali Ezzat, Mohammed Ali"
date: "December 30, 2017"
output: 
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) #data frame manipulation
library(ggplot2) #visualization
library(data.table) #reading table
library(magrittr) #code enhancement
library(scales) #for plot scaling
library(stringr) #string operations
library(quanteda) #text mining
library(gridExtra) # for plotting many plots in grids

```

# About
Mercari’s [challenge](https://www.kaggle.com/c/mercari-price-suggestion-challenge) is to build an algorithm that automatically suggests the right product prices. You’ll be provided user-inputted text descriptions of their products, including details like product category name, brand name, and item condition.

# Exploration
## Data Loading
* Load training data
```{r loadData, message = FALSE, warning = FALSE, tidy = TRUE}
train <- read.csv("Mercari-Price-Suggestion-Challenge/data/train.tsv", sep = "\t", stringsAsFactors = FALSE)
```

* Inspect structure
```{r inspectStructure}
glimpse(train)
```
* All about factors
It seems there are some columns are factors, let us convert them to the right format

```{r}
#item condition factor
train$item_condition_id <- as.factor(train$item_condition_id)
levels(train$item_condition_id)
table(train$item_condition_id)

#shupping
train$shipping <- as.factor(train$shipping)
levels(train$shipping)
table(train$shipping)
```
* Price summary

```{r}
summary(train$price)
```

## Feature relationships

```{r}
ggplot(train, aes(x = item_condition_id, y = log10(price))) +
  geom_boxplot()
```

```{r warning=FALSE}
ggplot(train, aes(x = shipping, y = log10(price))) +
  geom_boxplot()
```


```{r}

```

```{r}

```

```{r}
```

```{r}
```
# Wrangling
# Analysis
# Model Building
# Conclusion
# eBay acronyms
A-E
B&W: black and white

BC: back cover (usually used as a description for books)

BIN: Buy It Now

CIP: customer initiated payment

DOA: dead on arrival (an item that doesn't work or is broken when it's received)

DSR: detailed seller rating (additional Feedback ratings buyers can give sellers)

EST: Eastern Standard Time

EUC: excellent used condition

F-I
FAQ: frequently asked questions (a list of questions with answers.)

FB: Feedback

FC: fine condition

FOB: freight on board (usually means something has shipped)

FS: full screen (usually applied to a DVD or video format)

FVF: final value fee

G: good condition

GBP: Great Britain pounds

GU: gently used (item that has been used but shows little wear, accompanied by explanation of wear)

HP: home page

HTF: hard to find

HTML: hypertext markup language (the language used to create web pages)

IE: Internet Explorer

IM: instant messaging

INIT: initials

ISP: Internet service provider (a company that gives you access to the Internet)

J-M
JPG: JPEG (preferred file format for pictures on eBay, pronounced "jay-peg")

LTBX: letterbox (video format that recreates a widescreen image)

LTD: limited edition

MNT: mint or in perfect condition (a subjective term that doesn't necessarily mean new)

MIB: mint in box

MIJ: made in Japan

MIMB: mint in mint box

MIMP: mint in mint package

MIP: mint in package

MNB: mint no box

MOC: mint on card

MOMC: mint on mint card

MONMC: mint on near mint card

MWBT: mint with both tags

MWMT: mint with mint tags

N-P
NARU: not a registered user (also a suspended user)

NBW: never been worn

NC: no cover

NIB: new in box

NM: near mint

NOS: new old stock

NR: no reserve price (for an auction-style listing)

NRFB: never removed from box

NWT: new with tags

NWOB: new without box

NWOT: new without original tags

OEM: original equipment manufacturer

OOP: out of print

PST: Pacific Standard Time

Q-Z
RET: retired

SCR: scratch

S/O: sold out

Sig: signature


