---
title: "Mercari Price Suggestion Challenge"
author: "Ali Ezzat, Mohammed Ali"
date: "December 30, 2017"
output: 
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# About
Mercari’s [challenge](https://www.kaggle.com/c/mercari-price-suggestion-challenge) is to build an algorithm that automatically suggests the right product prices. You’ll be provided user-inputted text descriptions of their products, including details like product category name, brand name, and item condition.

# Exploration
* Load training data
```{r loadData, message = FALSE, warning = FALSE, tidy = TRUE}
train <- read.csv("Mercari-Price-Suggestion-Challenge/data/train.tsv", sep = "\t", stringsAsFactors = FALSE)
```

* Inspect structure
```{r inspectStructure}
glimpse(train)
```
* All about factors

It seems there are some columns are factors, let us convert them to the right format
 + item condition factor
```{r}
train$item_condition_id <- as.factor(train$item_condition_id)
levels(train$item_condition_id)
table(train$item_condition_id)
```
# Wrangling
# Analysis
# Model Building
# Conclusion
