---
title: "Damage Prediction Model"
author: "Mohammed Ali"
date: "August 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(useful)
library(h2o)
```

#Load Data

```{r}
buildings <- as_tibble(read.csv("buildings_final.csv"))
glimpse(buildings)
summary(buildings)
```


# Data Preperation

```{r}
buildings$X <- NULL #remove the index column
#buildings$damage_grade <- as.factor(buildings$damage_grade_id)
train <- subset(buildings, !is.na(buildings$buildings_all.damage_grade))
test <- subset(buildings, is.na(buildings$buildings_all.damage_grade))
summary(train)
```

# Feature Selection Approaches
Finding the most important predictor variables (of features) that explains major part of variance of the response variable is key to identify and build high performing models.

## Step-wise Regression method
```{r}
base.mod <- lm(damage_grade ~ 1 , data= train)  # base intercept only model
all.mod <- lm(damage_grade ~ . , data= train) # full model with all predictors
stepMod <- step(base.mod, scope = list(lower = base.mod, upper = all.mod), direction = "both", trace = 0, steps = 1000)  # perform step-wise algorithm
shortlistedVars <- names(unlist(stepMod[[1]])) # get the shortlisted variable.
shortlistedVars <- shortlistedVars[!shortlistedVars %in% "(Intercept)"]  # remove intercept
print(shortlistedVars)
summary(stepMod)
```

# Model Building using H2O

## Preparation

```{r}
localH2O <- h2o.init(nthreads = -1)
h2o.init()
train.h2o <- as.h2o(train)
test.h2o <- as.h2o(test)
```
Using column index, we need to identify variables to be used in modeling as follows:

```{r}
colnames(train.h2o)
```

```{r}
#dependent variable (damage_grade)
y.dep <- 81

#independent variables (from selection step)
x.indep <- c(2, 4:5, 16, 33:38, 55:56)
```


## Multiple Regression

```{r}
regression.model <- h2o.glm( y = y.dep, x = x.indep, training_frame = train.h2o, family = "multinomial")
h2o.performance(regression.model)
```

### Prediction
```{r}
predict.reg <- as.data.frame(h2o.predict(regression.model, test.h2o))
sub_reg <- data.frame(building_id = test$buildings_all.building_id, damage_grade = predict.reg$predict)
write.csv(sub_reg, file = "sub_reg.csv", quote = FALSE, row.names=FALSE)
```


## Random Forest
```{r}
 system.time(
rforest.model <- h2o.randomForest(y=y.dep, x=x.indep, training_frame = train.h2o, ntrees = 1000, mtries = 3, max_depth = 4, seed = 1122)
)
h2o.performance(rforest.model)
h2o.varimp(rforest.model)
```

### Prediction
```{r}
system.time(predict.rforest <- as.data.frame(h2o.predict(rforest.model, test.h2o)))
sub_rf <- data.frame(building_id = test$buildings_all.building_id, damage_grade = predict.rforest$predict)
write.csv(sub_rf, file = "sub_rf.csv", quote = FALSE, row.names=FALSE)
```

## GBM 

```{r}
system.time(
gbm.model <- h2o.gbm(y=y.dep, x=x.indep, training_frame = train.h2o, ntrees = 1000, max_depth = 4, learn_rate = 0.01, seed = 1122)
)
h2o.performance (gbm.model)
```

### Prediction
```{r}
predict.gbm <- as.data.frame(h2o.predict(gbm.model, test.h2o))
            
sub_gbm <- data.frame(building_id = test$buildings_all.building_id, damage_grade = predict.gbm$predict)
write.csv(sub_gbm, file = "sub_gbm.csv", quote = FALSE, row.names=FALSE)
```

## Deep Learning

```{r}
system.time(
             dlearning.model <- h2o.deeplearning(y = y.dep,
             x = x.indep,
             training_frame = train.h2o,
             epoch = 60,
             hidden = c(100,100),
             activation = "Rectifier",
             seed = 1122
             )
)
```

###Prediction

```{r}
predict.dl <- as.data.frame(h2o.predict(dlearning.model, test.h2o))
sub_dl <- data.frame(building_id = test$buildings_all.building_id, damage_grade = predict.dl$predict)
write.csv(sub_dl, file = "sub_dl.csv", quote = FALSE, row.names=FALSE)
```