---
title: "Startups Business Analytics"
author: "Mohammed Ali, Ali Ezzat"
date: "February 17, 2018"
output:
    html_document:
      toc : true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

# Objective
Our main goal here is to see how we can walk through a business case to analysis it and conclude useful results from it. Although the dataset here is dummy one but it has the same issues for any real dataset, except the huge size, like:

* A lot of predictors to comperhend and get usefl info from them.
* Data has a lot of missing, unclear items.
* Amiguate relation between the predictors

So let us begin our journey.

# Data Wrangling
First we will read the data using `read_csv` methof from `readr` packing within `tidyverse` package.

```{r read_data, message=FALSE}
startups <- read_csv("data/CAX_Startup_Data.csv")
```
So we have `r nrow(startups)` observations belongs to `r ncol(startups)` variables, one of them is the response variable that we would to designate later for predction and the rest are the predictors.

Let us see how the first 5 rows of the data looks like
```{r data_head}
head(startups, 5)
```

From that it seems we have the following notes:

* There are many missong values 
* Missing values are not only marked as `NA` but there are other values like `No Info` or just an empty string.
* Varaible are read as _characters_ datatype by default and would need to be probably converted to their _original_ data type(dates, factors, numeric)
* There are variables clearly will need special processing like `Short Description of company profile` , `Specialization of highest education` or `
Investors`
* There are issues in texual columns (upper and lower case, different format, etc..)
* Clearly `Dependent-Company Status` is the response variable.

So, the following are the steps we will conduct to clean our dataset

* Unify the way missing data are marked.
* Will convert the variables into their proper datatypes.
* Remove column with missed data over 40%.

## Setup Missing Data
```{r set_missing}
set_missing <- function(x) {
  # Replace 'No Info' with NA
  x[x == 'No Info'] <- NA
  # Replace empty string with NA
  x[x == ''] <- NA
  return(x)
}

startups <- map_df(startups, set_missing)
```

## Variables Correct type
### Factor Variables

Construct factor of factor variables and convert them

```{r convert_factors}
factor_cols <- c(2, 12, 16:17, 24, 26:65, 67, 71, 73, 75:87, 89:91, 93, 97, 100:101)
startups[, factor_cols] <- map_df(startups[, factor_cols], toupper)
startups[, factor_cols] <- map_df(startups[, factor_cols], as.factor)
```

Let us look at the `summary` for each variable to make sure that everything is ok.
```{r convert_factors_summary}
map(startups[, factor_cols], summary)
```

It seems that everything in order. However, there is a small note about some variables are about the *investors/owners/founders* not about the company itself which should not be there as certainly each person has his/her own qualifications and experience, but for sake of simplicity we will consider these varaibles are talking about all of these people at once as one unit.

### Numeric Variables
Construct factor of numeric variables and convert them

```{r convert_numeric, message=FALSE, warning=FALSE}
numeric_cols <- c(3:5,10,11,18:23,25,61,66,68:70,72,74,88,92,94:96,98,99,102:116)
startups[, numeric_cols] <- map_df(startups[, numeric_cols], as.numeric)
```

Let us look at the `summary` for each variable to make sure that everything is ok.
```{r convert_numeric_summary}
map(startups[, numeric_cols], summary)
```

It seems that seems that there are some outliers there, we will find out more we reach the EDA phase.

### Date Variables
Construct factor of date variables and convert them

```{r convert_date}
date_cols <- c(13:14)
startups[, date_cols] <- map_df(startups[, date_cols], mdy)
```

Let us look at the `summary` for each variable to make sure that everything is ok.
```{r convert_date_summary}
map(startups[, date_cols], summary)
```

## Remove polluted predictors
Now, as the variables in their proper data types, let us remove predictors with more than 40% missing data

```{r remove_poll_vars}
startups <- startups[colSums(is.na(startups))/nrow(startups) < .4]
dim(startups)
```
It seems we git rid off 3 variables, let us take another look at the data now
```{r top_5}
head(startups)
```

Much better.

## Textual Columns
These columns need special treatment to extract useful features from it. There could be no single method to be applied to all as the other types, we will treat each variable alone.

### Short Description of company profile
Until now I do not know what to do with it
```{r desc}
head(startups$`Short Description of company profile`, 10)
```

### Industry of company
It is a multi-level variable indicate the copmany industry, the levels are seprated by |.
```{r copm_indust}
head(startups$`Industry of company`, 10)
```

### Focus functions of company
It is a multi-value variable indicate the company foucs.
```{r comp_foc}
head(startups$`Focus functions of company`, 10)
```
The following are some notes about this field:
* values are either seprated by *,* or *&* and sometimes both.
* The characters case is not standarized.
* Missing functions

### Investors
It is a multi-value variable. Values are seprated by *|*, and sometimes there are missing values
```{r investors}
head(startups$Investors, 10)
```

# Exploratory Data Analysis

# Feature Engineering

# Pre-Modeling Processing
## Missing Values

## More Exploration

## Hypothesis Testing

## Predictor Selection
